SMT-Boosted Security Types for Low-Level MPC

### 2024-10-11 esop25/metalang.tex
checkpoint

The following Lemma establishes correctness of algorithmic confidentiality
type checking, and makes explicit that SMT checking is eliminated in the
judgement.
\begin{lemma}
  If $\atj{\toeq{\phi}}{R}{\ty}$ then $\eqj{R}{\eqs}{\phi}{\ty}$ for any $\eqs$.
\end{lemma}

@@

We equate types up to evaluation of subexpressions as defined in
Figure \ref{fig-notg}. Since expression evaluation is total we can
just evaluate types of closed instructions, which have closed types,
to obtain Hoare typings for instructions. In Figure \ref{fig-mtj} we
define the type derivation rules for $\metaprot$. In the
\TirName{Mesg} rule we use algorithmic confidentiality and integrity
type checking in a straightforward manner.  In the \TirName{Encode}
rule, we verify the hint given in the type annotation, and then use it
for type checking.  In the \TirName{App} rule we verify preconditions
of the given function type and instantiate the type and postconditions
with the given expression arguments.  The \TirName{Sig} rule we
specify how to verify $\Pi$ types for functions. Note that the values
$\mv_1,\ldots,\mv_n$ chosen to instantiate the function parameters are
chosen arbitrarily\cnote{Might need to be a little more careful about
  this}.

We allow the programmer to specify a type signature $\tsig$ that
maps functions in $\codebase$ to valid dependent Hoare
typings. Define:
\begin{definition}
  $\tsig$ is \emph{verified} iff $f : \tsig(f)$ is valid for all $f \in \dom(\tsig)$.
\end{definition}
In practice, the programmer only needs to specify pre-~and post-conditions,
since the rest of the type is reconstructed by type checking.

@@

>>>>

The following Lemma establishes correctness of algorithmic confidentiality
type checking, and makes explicit that SMT checking is eliminated in the
judgement.
\begin{lemma}
  \label{lemma-atj-sound}
  If $\atj{\toeq{\phi}}{R}{\ty}$ then $\eqj{R}{\eqs}{\phi}{\ty}$ for any $\eqs$.
\end{lemma}

@@

We equate types up to evaluation of subexpressions as defined in
Figure \ref{fig-notg}. Since expression evaluation is total we can
just evaluate types of closed instructions, which have closed types,
to obtain Hoare typings for instructions. In Figure \ref{fig-mtj} we
define the type derivation rules for $\metaprot$. In the
\TirName{Mesg} rule we use algorithmic confidentiality and integrity
type checking in a straightforward manner.  In the \TirName{Encode}
rule, we verify the hint given in the type annotation, and then use it
for type checking.  In the \TirName{App} rule we verify preconditions
of the given function type and instantiate the type and postconditions
with the given expression arguments.  The \TirName{Sig} rule we
specify how to verify $\Pi$ types for functions. Note that the values
$\mv_1,\ldots,\mv_n$ chosen to instantiate the function parameters are
chosen ``fresh'', meaning that any strings and client identifiers
occurring as substrings in $\mv_1,\ldots,\mv_n$ are unique and
not used in any program source code. This ensures generality
of verification as demonstrated in Lemma \ref{lemma-eqs-notg}.

We allow the programmer to specify a type signature $\tsig$ that
maps functions in $\codebase$ to valid dependent Hoare
typings. Define:
\begin{definition}
  $\tsig$ is \emph{verified} iff $f : \tsig(f)$ is valid for all $f \in \dom(\tsig)$.
\end{definition}
In practice, to define $\tsig$ the programmer only needs to specify
pre-~and postconditions, as we illustrate in Section
\ref{section-examples}, since the rest of the type is reconstructed by
type checking.

\subsection{Type Checking Soundness}

\begin{lemma}
  \label{lemma-fresh}
  Given $w$ and $\cid$ with $\fresh(w,\cid)$, if $\notg{\eqs}[w/y_1,\cid/y_n]$
  is not satisfiable, then there exists no $w',\cid'$ such that
  $\notg{\eqs}[w'/y_1,\cid'/y_n]$ is satisifiable.
\end{lemma}

\begin{proof}
  Let $\eqs = \notg{\eqs}[w/y_1,\cid/y_n]$ and $\eqs' =
  \notg{\eqs}[w'/y_1,\cid'/y_n]$. If $w',\cid'$ are fresh or chosen
  such that $\eqs'$ is an $\alpha$-renaming of $\eqs$, the result is
  immediate.  Otherwise, it must be the case that at least two
  distinct variables $x_1$ and $x_2$ in $\eqs$ are renamed as a single
  variable $x$ in $\eqs'$, given freshness of $w,\cid$. Let
  $\mathit{rename}$ be the function that maps variables in $\eqs$ to
  variables in $\eqs'$ such that $\mathit{rename}(\eqs) = \eqs'$. Let
  $\store'$ be a model of $\eqs'$, and define $\store$ such that:
  $$
  \store(x) \defeq \store'(\mathit{rename}(x)) 
  $$
  Then $\store$ must be a model of $\eqs$, which is a contradiction. \qed
\end{proof}

\begin{lemma}
  \label{lemma-eqs-notg}
  For all $\fresh(\mv_1,\ldots,\mv_n)$, $\notg{E_1}$, $\notg{E_2}$, assuming: 
  \begin{mathpar}
    \notg{\eqs_1}[\mv_1/y_1]\cdots[\mv_n/y_n]  \redx \eqs_1

    \notg{\eqs_2}[\mv_1/y_1]\cdots[\mv_n/y_n]  \redx \eqs_2

    E_1 \models E_2
  \end{mathpar}
  then there exists no $\mv_1',\ldots,\mv_n'$ such that:
    \begin{mathpar}
    \notg{\eqs_1}[\mv_1'/y_1]\cdots[\mv_n'/y_n]  \redx \eqs_1

    \notg{\eqs_2}[\mv_1'/y_1]\cdots[\mv_n'/y_n]  \redx \eqs_2

    E_1' \not\models E_2'
  \end{mathpar}  
\end{lemma}

\begin{proof}
  Given the assumptions, suppose that $E_1' \not\models E_2'$
  is an $\alpha$-renaming of $E_1 \models E_2$. In this
  case the consequence follows immediately. Otherwise,
  suppose on the contrary we can witness
  $E_1' \not\models E_2'$. By definition this means that
  $E_1' \wedge \neg E_2'$ is satisfiable
  but $E_1 \wedge \neg E_2$ is not, which is a contradiction
  by Lemma \ref{lemma-eqs-notg}. \qed 
\end{proof}

@@

\begin{proof}
  (Sketch) The result follows by induction on derivation of the
  main type judgement for $\instr$. The \TirName{Mesg} case follows
  by Lemma \ref{lemma-atj-sound}. The \TirName{Encode} and \TirName{App}
  cases follow by definition of the \TirName{Sig} rule and Lemma
  \ref{lemma-eqs-notg}. 
\end{proof}

### 2024-10-11 esop25/smt.tex
checkpoint

In Figure \ref{fig-eqs} we define a syntax of SMT-style constraints
$\eqs$, where constraint terms $\phi$ are similar to expressions $\be$
except the can ``cross party lines''. This turns out to be very useful
to express correctness properties-- for example, in the
Goldreich-Micali-Wigderson (GMW) protocol wire values in circuits are
represented by reconstructive shares \cite{evans2018pragmatic}.  If by
convention shares of values $n$ are represented by $\mesg{n}$ on
clients, then assuming two clients $\setit{1,2}$ the reconstructed
value can be expressed as $\mx{n}{1} + \mx{n}{2}$.  So, while summing
values across clients is disallowed in $\minicat$ protocols, we can
express properties of shared values via constraints.

\begin{fpfig}[t]{Syntax of constraints and memory application extended to
  SMT expressions}{fig-eqs}
$$
\begin{array}{rcl}
  \phi ::= x \mid \phi \fplus \phi \mid \phi \fminus \phi \mid \phi \ftimes \phi  \qquad
  \eqs ::= \phi \eop \phi \mid \eqs \wedge \eqs 
\end{array}
$$

\begin{mathpar}
  \store(\phi_1 \fplus \phi_2) = \cod{\store(\phi_1) \fplus \store(\phi_2)}

  \store(\phi_1 \ftimes \phi_2) = \cod{\store(\phi_1) \ftimes \store(\phi_2)}

  \store(\phi_1 \fminus \phi_2) = \cod{\store(\phi_1) \fminus \store(\phi_2)}
\end{mathpar}
\end{fpfig}

This language of constraints has a direct interpretation in the Finite
Fields theory of cvc5 \cite{SMFF}. We can leverage this to implement a
critical \emph{entailment} property, written $\eqs_1 \models \eqs_2$.
Our entailment relation is based on model satisfiability in a standard
sense, except we represent models as memories $\store$. To this end
we extend application of memories to terms $\store(\phi)$ in
Figure \ref{fig-eqs}, and we formally define entailment as follows.
\begin{definition}
  Given $\eqs$ of the form $\phi_1 \eop \phi_1' \wedge \cdots \wedge \phi_n \eop
  \phi_n'$ and $\store$, we say that $\store$ is a \emph{model} of
  $\eqs$, written $\store \models \eqs$, iff $\store(\phi_i) =
  \store(\phi_i')$ for all $1 \le i \le n$. We write $\eqs_1 \models
  \eqs_2$ iff  $\store \models E_1$ implies $\store \models
  E_2$ for all $\store$, and note this relation is reflexive and transitive.
\end{definition}

Given this Definition, the following Theorem is well-known and commonly
applied in SMT systems to implement our notion of entailment. 
\begin{theorem}
  $\eqs \models \phi \eop \phi'$ iff $\eqs \wedge \phi \not\eop \phi'$ is
  not satisfiable.
\end{theorem}

>>>>

We introduce the following syntax of SMT-style constraints
for $\minicat$:
$$
\begin{array}{rcl}
  \phi ::= x \mid \phi \fplus \phi \mid \phi \fminus \phi \mid \phi \ftimes \phi  \qquad
  \eqs ::= \phi \eop \phi \mid \eqs \wedge \eqs \mid \neg\eqs
\end{array}
$$
%
%In Figure \ref{fig-eqs} we define a syntax of SMT-style constraints
%$\eqs$, where
Note that constraint terms $\phi$ are similar to expressions $\be$
except that $\phi$ can ``cross party lines''. This turns out to be very useful
to express correctness properties-- for example, in the
Goldreich-Micali-Wigderson (GMW) protocol wire values in circuits are
represented by reconstructive shares \cite{evans2018pragmatic}.  If by
convention shares of values $n$ are represented by $\mesg{n}$ on
clients, then assuming two clients $\setit{1,2}$ the reconstructed
value can be expressed as $\mx{n}{1} + \mx{n}{2}$.  So, while summing
values across clients is disallowed in $\minicat$ protocols, we can
express properties of shared values via constraints.

%\begin{fpfig}[t]{Syntax of constraints}{fig-eqs}
%$$
%\begin{array}{rcl}
%  \phi ::= x \mid \phi \fplus \phi \mid \phi \fminus \phi \mid \phi \ftimes \phi  \qquad
%  \eqs ::= \phi \eop \phi \mid \eqs \wedge \eqs \mid \neg\eqs
%\end{array}
%$$
%
%\begin{mathpar}
%  \store(\phi_1 \fplus \phi_2) = \cod{\store(\phi_1) \fplus \store(\phi_2)}
%  
%  \store(\phi_1 \ftimes \phi_2) = \cod{\store(\phi_1) \ftimes \store(\phi_2)}
%  
%  \store(\phi_1 \fminus \phi_2) = \cod{\store(\phi_1) \fminus \store(\phi_2)}
%\end{mathpar}
%\end{fpfig}

This language of constraints has an obvious direct interpretation in the Finite
Fields theory of cvc5 \cite{SMFF}. We can leverage this to implement a
critical \emph{entailment} property, written $\eqs_1 \models \eqs_2$.
Our entailment relation is based on satisfiability in a standard
sense, where we represent models as memories $\store$ (mappings
from variables to values).
\begin{definition}
  We write $\store \models \eqs$ if $\store$ satisfies, aka is a model
  of, $\eqs$. We write $\eqs_1 \models
  \eqs_2$ iff  $\store \models E_1$ implies $\store \models
  E_2$ for all $\store$, and note this relation is reflexive and transitive.
\end{definition}
Given this Definition, the following Theorem is well-known and a fundamental
technique in SMT to implement our (common) notion of entailment. 
\begin{theorem}
  $\eqs \models \eqs'$ iff $\eqs \wedge \neg\eqs'$ is
  not satisfiable.
\end{theorem}
